---
title: "class15"
author: "Jason Patrick Bennett"
date: "May 22, 2018"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pathway Analysis

Lets first import our data:

```{r}
library(DESeq2)

metaFile <- "data/GSE37704_metadata.csv"
countFile <- "data/GSE37704_featurecounts.csv"
```

Format our Metadata:

```{r}
colData <- read.csv(metaFile, row.names = 1)
head(colData)
```

Now lets format our Countdata:

```{r}
countData <- read.csv(countFile, row.names = 1)
head(countData)
```

This "length" column in our data is out of place. We need to strip it from our data:

```{r}
countData <- as.matrix(countData[,-1])
head(countData)
```

Now we have the right number of columns, however, we see a lot of 0's included in the above dataset. We should get rid of those, as their inclusion would add nothing of value to our overall data:

```{r}
countData <- countData[rowSums(countData)>1, ]
head(countData)
```

Okay, now we are ready to create the DESeqDataSet object that will be used as input for our DESeq() function (Similar to the process done in the "class14" project):

```{r, cache=TRUE}
dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~condition)

dds <- DESeq(dds)
```

Lets check what we've created:

```{r}
dds
```

So now we have the required object; lets get results for our HoxA1 knockdown versus the control siRNA (labeled as "hoxa1_kd" and "control_sina" in our colData metaFile):

```{r}
res <- results(dds, contrast = c("condition", "hoxa1_kd", "control_sirna"))

# Reorder the results by p-value
res <- res[order(res$pvalue),]
summary(res)
```

We have results using the Ensembl annotations, however we will need to swap to Entrez gene IDs. Let's convert them now:

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

columns(org.Hs.eg.db)
```

And now the actual mapping:

```{r}
res$symbol <- mapIds(org.Hs.eg.db,
                     keys = row.names(res),
                     column = "SYMBOL",
                     keytype = "ENSEMBL",
                     multiVals = "first")

res$entrez <- mapIds(org.Hs.eg.db,
                     keys = row.names(res),
                     column = "ENTREZID",
                     keytype = "ENSEMBL",
                     multiVals = "first")

res$name <- mapIds(org.Hs.eg.db,
                   keys = row.names(res),
                   column = "GENENAME",
                   keytype = "ENSEMBL",
                   multiVals = "first")

head(res, 10)
```



## Pathway Analysis

Here we use the gage package for pathway analysis. Once we hae a list of enriched  pathways, we're going to use the pathview package to draw pathway diagrams, shading molecules in the pathway by their degree of up/down regulation.

### KEGG pathways

```{r}
# source("http://bioconductor.org/biocLite.R")

# biocLite( c("pathview", "gage", "gageData"))
```

Now lets load the packages:

```{r}
library(pathview)
library(gage)
library(gageData)

data(kegg.sets.hs)
data(sigmet.idx.hs)

kegg.sets.hs <- kegg.sets.hs[sigmet.idx.hs]
head(kegg.sets.hs, 3)
```

The gage() function requires a named vector of fold changes. We should generate that now:

```{r}
foldchanges <- res$log2FoldChange
names(foldchanges) <- res$entrez
head(foldchanges)
```

Now we can run the pathway analysis. We should check the function of the gage() function by looking at the ?gage help page.

```{r}
keggres <- gage(foldchanges, gsets = kegg.sets.hs, same.dir = TRUE)

attributes(keggres)
```

So we can see that we have a list object. It has three elements (greater, less, and stats). Lets use the $ to access the specific elements that we are interested in:

```{r}
head(keggres$greater)
```

```{r}
head(keggres$less)
```

These are data matrices with gene sets as rows sorted by p-value. Lets look at both up(greater), down(less), and statistics by using the head() with the lapply() function.

```{r}
lapply(keggres, head)
```

Now lets filter the results to extract the top 5 upregulated pathways, then extract just the IDs from those:

```{r}
pathways <- data.frame(id = rownames(keggres$greater), keggres$greater)
head(pathways)
```

Now lets use the pathview() function to make a pathway plot:

```{r}
pathview(gene.data = foldchanges, pathway.id = "hsa04110")
```

Lets generate a PDF graph of our results:

```{r}
pathview(gene.data = foldchanges, pathway.id = "hsa04110")
```

Now lets pull out the top 5 upregulated pathways, and then pull out only the IDs:

```{r, warning=FALSE}
keggrespathways <- rownames(keggres$greater)[1:5]

keggresids <- substr(keggrespathways, start = 1, stop = 8)
keggresids
```

Now lets pass these IDs in keggresids to the pathview() function to draw plots for all the top 5 pathways:

```{r, warning=FALSE}
pathview(gene.data = foldchanges, pathway.id = keggresids, species = "hsa")
```

We can check the generated graphs in our Files pane on the right.


## Gene Ontology (GO)

We can do a similar procedure with gene ontology:

```{r}
data(go.sets.hs)
data(go.subs.hs)
gobpsets <- go.sets.hs[go.subs.hs$BP]

gobpres <- gage(foldchanges, gsets = gobpsets, same.dir = TRUE)

lapply(gobpres, head)
```



## Reactome Pathways Analysis

To use the Reactome package, we first need to generate a plain text file containing the significant genes at the 0.05 level:

```{r}
sig_genes <- res[res$padj <= 0.05 & !is.na(res$padj), "symbol"]
print(paste("Total number of significant genes:", length(sig_genes)))
```

Now write this as a table to create the significant_genes.txt file. Then, upload this to the Reactome.org website to explore our data visually.

```{r}
write.table(sig_genes, file = "significant_genes.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)
```