---
title: "Structurla Bioinformatics"
author: "Jason Patrick Bennett"
date: "May 8, 2018"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Structural Bioinformatics

Download the CSV file

```{r}
pdb.stats <- read.csv("Data Export Summary.csv")
```

Let's find the totals and their associated percentages

```{r}
percent <- (pdb.stats$Total / sum(pdb.stats$Total) ) * 100
names(percent) <- pdb.stats$Experimental.Method
percent
```


## Working With Bio3D

Let's load the package

```{r}
# Already have this installed, so this is commented out (would need the first time)
#install.packages("bio3D")

# Load the package
library(bio3d)
```

List the functions in the package

```{r}
# Lists the functions available in the package
lbio3d()

# Shows an example of the specified function
example("plot.bio3d")
```

Let's read from PDB using our bio3D

```{r}
# Grabs the specified protein
pdb <- read.pdb("1hsg")
# Displays the information about the protein
pdb
```

Now lets write out a C-Alpha atom only PDB file so that we can view it in VMD

```{r}
inds <- atom.select(pdb, "calpha")
inds

# This line just confirms that we've selected the right portion of the data
pdb$atom[inds$atom,]

# Actually trim the pdb
ca.pdb <- trim.pdb(pdb, inds)

# Write to a new pdb with only the C-Alpha
pdb.selection <- write.pdb(ca.pdb, file = "CAlphaTrimmed.pdb")
```


## Multiple PDB Files

```{r}
ids <- c("1TND_B", "1AGR_A", "1TAG_A", "1GG2_A", "1KJY_A", "4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```

```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile = "C:/Users/thisi/Downloads/muscle.exe")
```


## Application on Adenylate Kinase

```{r}
aa <- get.seq("1ake_A")
```

Now do a Blast search. Much faster here than on BLAST website!
 - The cache = TRUE means that it caches our results for faster           subseqent searches
```{r, cache = TRUE}
b <- blast.pdb(aa)
```

Now plot our information to see similarity: the first 12 are indicated as being the most similar to our sample.

```{r}
hits <- plot(b)
```

Now lets look at the 12 that our BLAST identified

```{r}
(hits$pdb.id)
```

Now lets allign the structures:

```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile = "C:/Users/thisi/Downloads/muscle.exe")

ids <- basename.pdb(pdbs$id)

plot(pdbs, labels = ids)
```

Now lets use our PCA on the data:

```{r}
# Find invariant core
core <- core.find(pdbs)

# Superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)

# Perform pca
pc.xray <- pca(pdbs)
```

