---
title: "UKFoods"
author: "Jason Patrick Bennett"
date: "May 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## UK Foods Report

Read the file into R
```{r}
data <- read.csv("https://bioboot.github.io/bimm143_S18/class-material/UK_foods.csv")

head(data)
```

Head shows that our row names are included as a column, which we don't want

Set the first column of data to be the rownames of data
```{r}
rownames(data) <- data[,1]

# Take out the first line of data so that the rownames are no longer included
data <- data[,-1]

head(data)
```


Display our data as a rendered table using the Knitr package
```{r}
knitr::kable(data, caption = "The full UK foods data table")
```

The raw numbers don't show any obvious trends, so lets apply a heatmap to the data
```{r}
heatmap(as.matrix(data))
```

This helps emphasize the differences between countries, however it doesn't quantify the data

Let's use PCA to help us evaluate more closely

# PCA

Use prcomp() to perform a pca analysis
```{r}
# Remeber to transpose the dataset before using prcomp if your observations are columns and not rows
pca <- prcomp(t(data))

summary(pca)
```

By analyzing the above summary, we can see that our PCA has revealed that PC1 is responsible for 67% of the Perceived Variance!

Together with PC2, 96.5% of the Perceived Variance is explained. Clearly these are areas we want to explore more

Lets plot PC1 vs PC2
```{r}
plot(pca$x[,1], pca$x[,2], xlab = "PC1", ylab = "PC2", xlim = c(-270, 500) )

# Places the country names within the plot instead of simple points
text(pca$x[,1], pca$x[,2], colnames(data))
```

Now lets access the PCE for each PC
```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100)

# This displays the PVE for each PC from PC1 down through our list
v
```

We can visualize this using a plot of the variances (the Eigenvalues) with respect to the Principle Component number (the Eigenvector numbers)
```{r}
barplot(v, xlab = "Principle Component", ylab = "Percent Variation")
```

To show how much of the PVE is accounted for in our Principle Componenets, we can take the cumulative sum of our PVEs
```{r}
cumsum(v)
```

To visualize the impact each of the original variables had on the PVE, we can look at the **Loading Scores** accessed by inspection of the pca$rotation componenet
```{r}
par(mar=c(10, 4, 4, 2))
barplot( pca$rotation[,1], las=2 )
```

This shows us that the variables that push a country furthest to the positive side of the graph are those who consume more Fresh Potatoes and Soft Drinks. North Ireland was a major consumer of both, hence why North Ireland was found so strongly on the positive side.

We can also see that the variables that produced the largest negative scores were Fresh Fruit and Alcoholic Drinks. This explains the countries that are major consumers of these variables and find themselves to the negative side of the graph

We can see these ideas more easily represented in a biplot: This shows the countries and their positions along with the relative weights of each variable with their associated vectors. The magnitude of the vectors correlates with their weight within PC1.
```{r}
biplot(pca)
```

## PAC on RNA-seq Data

First lets generate some sample data where:

Columns: Observations
Rows: Measurements
```{r}
# Just for the sake of the example, here's some made up data...
data.matrix <- matrix(nrow=100, ncol=10)

# Lets label the rows gene1, gene2 etc. to gene100
rownames(data.matrix) <- paste("gene", 1:100, sep="")

# Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
#   and the last 5 ko1, ko2 etc. to ko5 (for "knock-out") 
colnames(data.matrix) <- c(
  paste("wt", 1:5, sep=""),
  paste("ko", 1:5, sep=""))

# Now fill in some fake read counts
for (i in 1:100) {
  wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
  ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 
  data.matrix[i,] <- c(wt.values, ko.values)
}

head(data.matrix)

# Citation: Bary J. Grant, BIMM143, 2018
```

Now that we have our data, look at the matrix: Our Samples(OBSERVATIONS) are Columns and our Genes are Rows. Remember that our prcomp() wants the Observations as Rows.

We need to transpose our matrix before using prcomp() on it.

Also, the values vary wildly, so make sure to set SCALE = TRUE.
```{r}
pca <- prcomp(t(data.matrix), scale = TRUE)

# Plot our PCA looking at PC1 vs. PC2
plot(pca$x[,1], pca$x[,2], xlab = "PC1", ylab = "PC2")
```

Lets calculate how much variance each PC is responsible for (the Perceived Variance Explained)
```{r}
pca.var <- pca$sdev^2

pca.var.per <- round(pca.var/sum(pca.var) * 100, 1)
pca.var.per
```

Now use this data to generate our standard Scree-Plot
```{r}
barplot(pca.var.per, main = "Scree Plot", xlab = "Principle Component", ylab = "Percent Variation")
```

Now lets add names and colors to our original PCA plot to make it more useful
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(data.matrix)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"

plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))

text(pca$x[,1], pca$x[,2], labels = colnames(data.matrix), pos=c(rep(4,5), rep(2,5)))

#Citation: Barry J Grant, BIMM143, 2018
```

Finally let's show the genes in order of those that contribute the most to PC1 (Loading Scores)
```{r}
loading_scores <- pca$rotation[,1]

## Find the top 10 measurements (genes) that contribute
## most to pc1 in eiter direction (+ or -)
gene_scores <- abs(loading_scores) 
gene_score_ranked <- sort(gene_scores, decreasing=TRUE)

## show the names of the top 10 genes
top_10_genes <- names(gene_score_ranked[1:10])
top_10_genes
```

