---
title: "Drug Interactions"
author: "Jason Patrick Bennett"
date: "May 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Drug Discovery

First load the Bio3D package and download our desired structure

```{r}
# Load the package
library(bio3d)

# Retrieve the HIV-1 Protease PDB file
file.name <- get.pdb("1hsg")

# Save the file to the variable 'hiv'
hiv <- read.pdb(file.name)

# Quick look at the object
hiv
```

**Question 1: What is the name of the two non-protein resid values?**
  Water (HOH) and MK1
  
**Question 2: What does resid correspond to and how would we get a listing of all residue values in this structure?**
  The position within the 1HSG where these things exist. You could get   a list of all the residue values by using: "hiv$atom$resid"
  
  
Now lets trim the file to make it easier to use for our data analysis

```{r}
# Select only the protein
prot <- trim.pdb(hiv, "protein")

# Select only the ligand (or the drug)
lig <- trim.pdb(hiv, "ligand")
```


And lets take these trimmed sections and create new pdb files from them

```{r}
# Create the protein file
write.pdb(prot, file = "1hsg_protein.pdb")

# Create the ligand file
write.pdb(lig, "1hsg_ligand.pdb")
```


Now use AutoDock to visualize the molecule

**Question 3: Can we see 