---
title: "Drug Interactions"
author: "Jason Patrick Bennett"
date: "May 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Drug Discovery

First load the Bio3D package and download our desired structure

```{r}
# Load the package
library(bio3d)

# Retrieve the HIV-1 Protease PDB file
file.name <- get.pdb("1hsg")

# Save the file to the variable 'hiv'
hiv <- read.pdb(file.name)

# Quick look at the object
hiv
```

**Question 1: What is the name of the two non-protein resid values?**
  Water (HOH) and MK1
  
**Question 2: What does resid correspond to and how would we get a listing of all residue values in this structure?**
  The position within the 1HSG where these things exist. You could get   a list of all the residue values by using: "hiv$atom$resid"
  
  
Now lets trim the file to make it easier to use for our data analysis

```{r}
# Select only the protein
prot <- trim.pdb(hiv, "protein")

# Select only the ligand (or the drug)
lig <- trim.pdb(hiv, "ligand")
```


And lets take these trimmed sections and create new pdb files from them

```{r}
# Create the protein file
write.pdb(prot, file = "1hsg_protein.pdb")

# Create the ligand file
write.pdb(lig, "1hsg_ligand.pdb")
```


Now use AutoDock to visualize the molecule

**Question 3: Can we see **

## TODO: FINISH THIS






Now lets inspect the docking results

```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)

write.pdb(res, "results.pdb")
```


Pulling this up in VMD with the 1hsg.pdb file, we can see both the protein and our results superimposed on eachother. This allows us to *qualitatively* determine if our assumptions were correct.

---- Use VMD to compare the two ----

**Question 4: How good do the docks look qualitatively? Is the Crystal Binding Mode reproduced? Is it the best conformation according to AutoDock Vina?**
  They look good. ? ?

They look good, but lets determine *quantitatively* if they really are good.

```{r}
ori <- read.pdb("ligand.pdbqt")

rmsd(ori, res)
```


**Question 5: How good are the docks quantitatively? Is the Crystal Binding Mode reproduced within 1 Angstrom RMSD for all atoms?**
  Our first dock is within 1 Angstrom, however the rest fall short.     Therefore, our first dock is the best.
  
**Question 6: How would you determine the RMSD for heavy atoms only (i.e. non Hydrogren atoms)?**
  
  
```{r}
inds.res <- atom.select(res, string = "noh")

rmsd(lig, res$xyz[,inds.res$xyz])
```

