---
title: "class13"
author: "Jason Patrick Bennett"
date: "May 15, 2018"
output:
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Identifying SNP's in a Population

Lets analyze SNP's from the Mexican-American population in Los Angeles:

```{r}
genotype <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

Now lets look at a table of the data:

```{r}
table(genotype)
```

Isolate the data that we need from the table:

```{r}
genotype.summary <- table(genotype[,2])
genotype.summary
```

Find the proportion of each genotype:

```{r}
genotype.summary/nrow(genotype) * 100
```




## RNA Seq

Look more closely at sample: HG00109 (M) with genotype G|G

**Jetstream IP: 129.114.16.25**

Side-note on Fastq quality scores

```{r}
#install.packages("seqinr")
#install.packages("gtools")

# Sample read
library(seqinr)
library(gtools)
phred <- asc( s2c("DDDDCDEDCDDDDBBDDDCC@")) - 33
phred

# The Probability of the base being incorrect
prob <- 10**(-phred/10)
prob
```



## RNA-Seq Population Scale Analysis

Now lets analyze a text file of the reads from RNA-Seq (~230 samples or People)

```{r}
expr <- read.table("https://bioboot.github.io/bimm143_S18/class-material/
rs8067378_ENSG00000172057.6.txt")
```

Look at data:

```{r}
table(expr$geno)
```

How many patients?

```{r}
nrow(expr)
```

Summary of the data might be insightful

```{r}
summary(expr)
summary( expr$exp[ (expr$geno == "G/G")])
summary( expr$exp[ (expr$geno == "A/A")])
summary( expr$exp[ (expr$geno == "A/G")])
```


Now lets plot the data

```{r}
p <- boxplot(exp ~ geno, data = expr, notch = TRUE)
```

```{r}
plot(expr$geno, expr$exp)
```


Now lets try GGPlot:

```{r}
library(ggplot2)
ggplot(expr, aes(geno, exp)) + geom_boxplot()
```

```{r}
ggplot(expr, aes(exp, fill = geno)) + geom_density(alpha = 0.2)
```

And finally, the fanciest graph!

```{r}
ggplot(expr, aes(geno, exp, fill=geno)) + 
  geom_boxplot(notch=TRUE, outlier.shape = NA) + 
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.4)
```
