---
title: "Introduction to R Functions"
author: "Jason Patrick Bennett"
date: "June 1, 2018"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Functions

Lets first try a standard add function:

```{r}
# Standard add function
add <- function(x, y=1) {
  # Sum x and y inputs
  x + y
}
```

And now, a function to rescale data:

```{r}
# Rescaling function
rescale <- function(x) {
  rng <- range(x)#, na.rm = TRUE)
  rescaled <- (x - rng[1]) / (rng[2] - rng[1])
  return(rescaled)
}
```

Now we'll try a more advanced rescale function that also includes a plot depending on a logical parameter passed to the function:

```{r}
# Second rescaling function
rescale2 <- function(x, na.rm = TRUE, plot = FALSE) {
  if(na.rm) {
    rng <- range(x, na.rm = TRUE)
  } else {
    rng <- range(x)
  }
  
  answer <- (x - rng[1]) / (rng[2] - rng[1])
  
  if(plot) {
    plot(answer, typ = "b", lwd = 4)
  }

  return(answer)
}
```

Lets test the first function now:

```{r}
# Test on a small example where you know the answer
rescale(1:10)
```

```{r}
# Displays NA for all; Not optimal
rescale( c(1, 2, NA, 3, 10) )
```

This second rescale result was not exactly what we were looking for. What can we do to improve this?

  - Remove NA from the dataset before trying to rescale! (Uncomment the na.rm = TRUE      segment of the rescale function).

```{r}
rescale2( c(1, 2, NA, 3, 10))
```