---
title: "Bioinformatics Class 7"
author: "Jason Patrick Bennett"
date: "April 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions: *Part Deux*

Here I am going to revsit our function from class 6: **rescale**.

```{r}
source("http://tinyurl.com/rescale-R")
```

Lets see if we can use this function!

```{r}
rescale(1:10)
```

Looks good! Lets break it!

```{r}
#rescale( c(1:10, "break") )
```

Lets try the new **rescale2()** function.

```{r}
#rescale2( c(1:10, "brokenButBetter") )
```

## Write a NA checking function

Here we write a new function to check for NAs in two inputs

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```


```{r}
is.na(x)
```


```{r}
which (is.na(x) )
```


```{r}
sum ( is.na(x) )
```


```{r}
sum ( is.na(x) & is.na(y) )
```


```{r}
which ( is.na(x) & is.na(y) )
```


```{r}
both_na(x, y)
```


```{r}
x <- c(NA, NA, NA)
y1 <- c(1, NA, NA)
y2 <- c(1, NA, NA, NA)

#both_na(x, y2)
```


```{r}
#both_na2(x, y2)
```


```{r}
both_na3(x, y1)
```


## Another function example: gene intersection

```{r}
df1
df2

x <- df1$IDs
y <- df2$IDs
```


```{r}
x
y
```


```{r}
#intersect(x, y)

x %in% y
```


```{r}
y %in% x
```


Now we can access the genes we want with these indices!


```{r}
x[x %in% y]
y[y %in% x]
```


Can make these columns of the same object with **cbind()**


```{r}
cbind( x[ x %in% y], y[ y %in% x] )
```


```{r}
gene_intersect(x, y)
```


```{r}
gene_intersect2(df1, df2)
```


Lets try the **merge()** function for this job


```{r}
merge(df1, df2, by="IDs")
```

